---
title: "boston airbnb maps"
author: "Connor King"
date: "2023-06-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mapping Price

Here is the beginning code from the exploratory data analysis file that loads in the cleaning listings dataset and creates a box plot using log prices.

```{r, include = FALSE}
library(tidyverse)
library(plotly)
library(mapboxapi)
library(shiny)
library(sf)

listings <- read_csv("listings_2023.csv")

listings %>% 
  filter(price == 0) %>% 
  select(name, room_type, price)

boston_airbnb <-
  listings %>% 
  filter(price > 0 & price <= 9999) %>% 
  mutate(log_price = log(price, 10)) 

boston_airbnb$neighbourhood[boston_airbnb$neighbourhood == 'Longwood Medical Area'] = 'Longwood'

lvls <- 
  boston_airbnb %>% 
  group_by(neighbourhood) %>% 
  summarise(m = median(price)) %>% 
  arrange(m) %>% 
  pull(neighbourhood)

p_box <-
  plot_ly(
    boston_airbnb,
    x = ~factor(neighbourhood, lvls),
    y = ~price,
    type = "box",
    showlegend = FALSE,
    name = ""
  ) %>% 
  layout(
    yaxis = list(type = "log", title = "log(price)"),
    xaxis = list(title = "", tickangle = -35)
  )
```

Here is the `plotly` mapbox with just listings mapped.

```{r}

Sys.setenv("MAPBOX_TOKEN" = "pk.eyJ1IjoiYm9zdG9uY29ubm9yMTEiLCJhIjoiY2xncXIya2VxMGc1cTNmc2I3NjFoY2NkMyJ9.fcKb-W66WPlzv4oOx6ZC4A")

p <- 
  plot_mapbox(boston_airbnb, width = 1200, height = 900) %>% 
  add_markers(
    x = ~longitude,
    y = ~latitude,
    color = ~log_price,
    name  = "Log (base 10) of price",
    text = 
      ~paste(
        name,
        "\nRoom type:", room_type,
        "\nPrice: ", price,
        "\nNeighborhood: ", neighbourhood,
        "\nMinimum nights: ", minimum_nights
      ),
    hoverinfo = "text"
  ) %>% 
  layout(
    mapbox = 
      list(
        center = list(lat = 42.32, lon = -71.1),
        zoom = 9.5,
        style = "dark"
      )
  )

p
```


## Adding T stations and T lines

I will insert a `GPX` formatted file of MBTA stations from  http://erikdemaine.org/maps/mbta/ which includes all the waypoints of the MBTA stations as well as routes of rapid transit lines.

I will use `read_GPX()` function from `tmaptools` package which reads `GPX` files into `sf` objects in R.

```{r}
library(tmaptools)
mbta <- read_GPX("mbta.gpx")

stations <-
  mbta$waypoints %>%
    filter(grepl('Red Line|Green Line|Blue Line|Orange Line', type))

T_lines <-
  mbta$tracks %>%
    filter(grepl('Red Line|Green Line|Blue Line|Orange Line', name))

  (p %>% 
  add_sf(
    data = stations,
    inherit = FALSE,
    name = "MBTA T stations",
    text = ~name,
    hoverinfo = "text"
  ) %>% 
  add_sf(
    data = T_lines,
    text = ~name,
    hoverinfo = "text",
    name = "MBTA T lines"
  )
)

```

Now we will change the lines according to color. 

```{r}
T_lines %>% 
  select(name)
```

As shown, there are multiple lines with a respective color. Thus I will create a function that filters through the `name` column and adds an `sf` layer.

```{r}
add_MBTA_line <- function(p, line_color) {
  res <-
    p %>% 
      add_sf(
        data = T_lines %>% filter(grepl(line_color, name, ignore.case = TRUE)),
        color = ~I(line_color),
        text = ~name,
        hoverinfo = "text",
        name = paste0(line_color, " line")
      )
  return(res)  
}

p_lines <- 
  p %>% 
    add_sf(
      data = stations,
      name = "MBTA T stations",
      text = ~paste0(name, " (", type, ")"),
      hoverinfo = "text",
      color = I("pink"),
      size = I(30)
    ) %>%
    add_MBTA_line("red") %>%
    add_MBTA_line("green") %>%
    add_MBTA_line("orange") %>%
    add_MBTA_line("blue") %>% 
    colorbar(title = "log(price)")

p_lines

```

## Map with Neighborhood Boundaries

I will use a `KML` formatted file from https://data.boston.gov/dataset/boston-neighborhoods that contains the boundaries of all of Boston's neighborhoods.

To read the `KML` data into `sf`, we will use `st_read` from the `sf` package.

```{r}
boston_neighborhoods <- sf::st_read("Boston_Neighborhoods.kml")

p1 <- 
  p %>%
    add_sf(
      inherit = FALSE,
      data = boston_neighborhoods,
      fill = "",
      name = "Neighborhoods Boundaries",
      text = ~Name,
      hoverinfo = "text",
      color = I("azure4")
    ) %>% 
    colorbar(title = "log(price)")

pp <- subplot(p_box, p1, nrows = 2, heights = c(0.2, 0.8), margin = 0.1)
pp

```


## Additional Mobility and Accessbility Infrastructure

Now I will incorporate Bluebikes stations and Police Stations. 

```{r}
police <- read_csv("Boston_Police_Stations.csv")
bikes <- read_csv("Blue_Bike_Stations.csv")
meters <- read_csv("Parking_Meters.csv")
hospitals <- read_csv("Hospitals.csv")
ev <- read_csv("Charging_Stations.csv")
open_space <- sf::st_read("Open_Space.kml")


meters <- meters %>%
  separate(PAY_POLICY, into = c("pay_time", "pay_days", "pay_rate", "pay_duration"), sep = " ", extra = "merge") %>% 
  mutate(pay_duration = as.numeric(pay_duration)/60)
```


```{r}

p_h <- p %>% 
  add_trace(
    data = hospitals,
    x = ~Longitude,
    y = ~Latitude,
    mode = 'markers',
    name = "Hospitals",
    text = ~paste(Name,
                  "\nAddress: ", Address),
    hoverinfo = "text",
    marker = list(
      color = 'magenta', 
      size = 7)
  ) %>% 
  add_trace(
    data = bikes,
    x = ~Longitude,
    y = ~Latitude,
    opacity = 0.5,
    mode = 'markers',
    name = "Bike Stations",
    text = ~paste("Station Number: ", Number,
                  "\nAddress: ", Name,
                  "\nTotal Docks: ", Total_docks),
    hoverinfo = "text",
    marker = list(
      color = 'navy', size = 5)
  ) %>% 
  add_trace(
    data = police,
    x = ~CENTROIDX,
    y = ~CENTROIDY,
    mode = 'markers',
    name = "Police Stations",
    text = ~paste("Address: ", ADDRESS),
    hoverinfo = "text",
    marker = list(color = 'red', size = 10)
  ) %>% 
  add_trace(
    data = ev,
    x = ~Longitude,
    y = ~Latitude,
    mode = 'markers',
    name = 'EV Charging Stations',
    text = ~paste(
      Station_Name,
      "\nAddress: ", Street_Address,
      "\nNetwork ", ifelse(is.na(EV_Network), "Not Available", EV_Network),
      "\nConnector: ", EV_Connector_Types
    ),
    hoverinfo = "text",
    marker = list(color = "orange", size = 7)
  ) %>% 
  add_trace(
    data = meters,
    x = ~LONGITUDE,
    y = ~LATITUDE,
    opacity = 0.5,
    mode = 'markers',
    name = 'Parking Meters',
    text = ~paste(
      "Payment Days: ", pay_days,
      "\nPayment Times: ", pay_time,
      "\nRate (per hour): ", pay_rate,
      "\nDuration (hours): ", pay_duration
    ),
    hoverinfo = "text",
    marker = list(color = "#8c564b", size = 7)
  ) %>% 
  add_sf(
      inherit = FALSE,
      data = open_space,
      name = "Open Spcae",
      text = ~Name,
      hoverinfo = "text",
      color = I("darkgreen")
  )

p_h
```

By visualizing Bluebikes station locations alongside Airbnb listings, potential renters can easily identify listings with convenient access to bike-sharing facilities. This could be a crucial deciding factor for those prioritizing active transportation, and also provides a glimpse into the infrastructure of the city.

The location of police stations can be a proxy for the security and safety infrastructure of the city. By visualizing police station locations, I offer viewers an opportunity to assess the accessibility of police services from different listings. While it's not a direct measure of a neighborhood's safety, proximity to a police station might be an important consideration for some renters.

#### Sources

Bluebikes: https://data.boston.gov/dataset/blue-bike-stations
Police Stations: https://data.boston.gov/dataset/boston-police-stations

### Choropleth Map 

Here's a Choropleth Map using `ggplot` that visualizes the average price for each neighborhood.

```{r, fig.width= 10, fig.height= 10}

library(viridis)

neigh_means <- boston_airbnb %>% 
  group_by(neighbourhood) %>% 
  summarise(mean = mean(price)) %>% 
  rename(Name = neighbourhood) #to match boston_neighborhoods

df_neigh <- na.omit(left_join(boston_neighborhoods, neigh_means, by = "Name"))

g_choro <- ggplot() +
  geom_sf(data = df_neigh, aes(fill = mean), color = NA) +
  geom_sf_text(data = df_neigh, aes(label = Name), size = 2, color = "red") +
  scale_fill_gradientn(colors = viridis(100)) +
  theme_minimal()

g_choro

p_choro <- ggplotly(g_choro)

ggsave("neigh choro.png", plot = g_choro, dpi = 300, width = 12, height = 12)
```

























