---
title: "airbnb eda"
author: "Connor King"
date: "2023-06-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

links:
listings http://insideairbnb.com/get-the-data/

## Dataset

```{r}
library(tidyverse)
library(plotly)
#listings from march 19 2023
listings <- read_csv("listings_2023.csv")
  
Sys.setenv("MAPBOX_TOKEN" = "pk.eyJ1IjoiYm9zdG9uY29ubm9yMTEiLCJhIjoiY2xncXIya2VxMGc1cTNmc2I3NjFoY2NkMyJ9.fcKb-W66WPlzv4oOx6ZC4A")
```

```{r}
summary(listings$price)
```
```{r}
plot_ly(listings, y = ~price, type = "box")
```

```{r}
plot_ly(listings, x = ~price, type = "histogram")
```

As the boxplot shows, there appears to be some heavy outliers especially the listings around 10k.


```{r}
head(listings$price[order(listings$price)], )
```
Given the context, the two listings with a price of $0 do not make sense.

```{r}
tail(listings$price[order(listings$price)], 20)
```
```{r}
sum(listings$price > 2000)

```
There are 7 units priced over $2,000.

```{r}
listings %>% 
  filter(price > 2000) %>% 
  select(name, room_type, price)

```

The listing with the price of $10,000 seems to be an error as it's a private room with only a double bed. The other high priced listings however, are certainly plausible.

```{r}
listings %>% 
  filter(price == 0) %>% 
  select(name, room_type, price)
```
These listings also appear to be errors.

Thus I will remove the listings with a price of $0 and the listing with a price of 10,000.

```{r}
boston_airbnb <-
  listings %>% 
  filter(price > 0 & price <= 9999) %>% 
  mutate(log_price = log(price, 10)) #for color scale in the maps
```

Now we will create a boxplot that demonstrates the effect of neighborhood on price (arranged in increasing order of their median price).

```{r}
#for arrangement
lvls <- 
  boston_airbnb %>% 
  group_by(neighbourhood) %>% 
  summarise(m = median(price)) %>% 
  arrange(m) %>% 
  pull(neighbourhood)

p_box <-
  plot_ly(
    boston_airbnb,
    x = ~factor(neighbourhood, lvls),
    y = ~price,
    type = "box",
    showlegend = FALSE,
    name = ""
  ) %>% 
  layout(
    yaxis = list(type = "log", title = "log(price)"),
    xaxis = list(title = "", tickangle = -35)
  )

p_box
```

